Iterative Monomorphisation

### 2024-02-21 iterative_monomorphisation.tex
more

\begin{definition}
A (\emph{type}) \emph{substitution} is a partial function mapping a finite number of type variables to corresponding types. Substitutions are written as
$\sigma = \{\alpha_1\mapsto\tau_1, \dots, \alpha_n\mapsto\tau_n\}$. They are assumed to be lifted to formulae; thus, $\sigma(\varphi)$ yields the variant of $\varphi$ in which each $\alpha_i$ is replaced by $\tau_i$.

Given \(\tau\) and \(\upsilon\) two substitutions, \(\tau \circ \upsilon\) is the successive application of \(\upsilon\) then \(\tau\).
\end{definition}

@@

%We start with a high level description of the iterative monomorphisation algorithm.
The iterative monomorphisation algorithm takes a polymorphic problem as input and returns a monomorphic problem. It operates by applying an arbitrary number of iterations. Each iteration takes a polymorphic problem as argument and returns a problem with additional partially instantiated formulae. A single iteration consists of a collection phase (Section~\ref{ssec:collection}) and a processing phase (Section~\ref{ssec:processing}).

@@

It is important to parametrise \(P\) with \(\varphi\) because type variables are quantified at the formula level. The formula indicates the scope of [the?] type variables. This is not necessary for \(M\) since all the types it contains are monomorphic.

@@

Each iteration follows these steps:
\begin{enumerate}

   \item The maps \(M\) and \(P\) are initialised with respect to \(\Phi\) as per the collection phase.

   \item An empty set of formulae \(\Psi\) is created.

   \item For each formula \(\varphi \in \Phi\) and for each symbol \(f\) occuring in \(\varphi\):

      An empty set of substitutions \(\Sigma_{\varphi, f}\) is created, then for each tuple \((\tau_1, \dots, \tau_n) \in  P(\varphi, f)\) and \((\upsilon_1, \dots, \upsilon_n) \in M(f)\):

      For all \(i\), we match \(\tau_i\) against \(\upsilon_i\), possibly yielding the substitution \(\sigma_i\). If all \(n\) matchings are successful and the substitutions are two by two compatible, \(\sigma_1 \circ \dots \circ \sigma_n\) is added to \(\Sigma_{\varphi, f}\).

   \item For each formula \(\varphi \in \Phi\), for each symbol \(f\) in \(\varphi\) and for each substitution \(\sigma \in \Sigma_{\varphi, f}\), the formula \(\sigma(\varphi)\) is added to \(\Psi\).

   \item \(\Phi \cup \Psi\) is returned.

@@

An arbitrary number of iterations can be executed, using the output of one iteration as the input for the next. The final step requires filtering all monomorphic formulae returned by the last iteration.

>>>>

\begin{definition}
A (\emph{type}) \emph{substitution} is a partial function mapping a finite number of type variables to corresponding types. Substitutions are written as
$\sigma = \{\alpha_1\mapsto\tau_1, \dots, \alpha_n\mapsto\tau_n\}$. They are assumed to be lifted to formulae; thus, $\sigma(\varphi)$ yields the variant of $\varphi$ in which each $\alpha_i$ is replaced by $\tau_i$.
Given two substitutions \(\tau, \upsilon\), the successive application of \(\tau\) and \(\upsilon\) is denoted by \(\upsilon \circ \tau\).
\end{definition}

@@

%We start with a high level description of the iterative monomorphisation algorithm.
The iterative monomorphisation algorithm takes a polymorphic problem as input and returns a monomorphic problem. It operates by applying an arbitrary number of iterations. Each iteration takes a polymorphic problem as argument and returns a problem with additional partially instantiated formulae. A single iteration consists of a collection phase (Section~\ref{ssec:collection}) and a processing phase (Section~\ref{ssec:processing}). Once the iterations are completed, a final step filters out all non-monomorphic formulae returned by the last iteration.

@@

It is important to parametrise \(P\) with \(\varphi\) because type variables are quantified at the formula level. The formula indicates the scope of type variables. This is not necessary for \(M\) since all the types it contains are monomorphic.

@@

Once the maps \(M\) and \(P\) are initialised, each iteration performs the following processing steps:

\begin{enumerate}

   \item Create an empty set of formulae \(\Phi'\).

   \item For each formula \(\varphi \in \Phi\) and for each symbol \(f\) occuring in \(\varphi\):
   \begin{enumerate}
    \item[2.1.] For each tuple \((\tau_1, \dots, \tau_n) \in  M(f)\)        and each tuple \((\upsilon_1, \dots, \upsilon_n) \in P(\varphi, f)\),
     for each \(i\), match \(\upsilon_i\) against \(\tau_i\), yielding the substitution \(\sigma_i\) in case of success.

    \item[2.2.] If all \(n\) matchings are successful and the substitutions \(\sigma_i\) are pairwise compatible,
add the formula \((\sigma_1 \circ \dots \circ \sigma_n)(\varphi)\) to \(\Phi'\).
   \end{enumerate}

   \item Return \(\Phi \cup \Phi'\).

@@

An arbitrary number of iterations can be executed, using the output of one iteration as the input of the next.
